---
title: "validateScenarios"
author: "Pascal Weigmann"
date: "2024-03-21"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{validateScenarios}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(piamValidation)
```

# Overview

The function `validateScenarios()` performs validation checks on IAM scenario data based on thresholds provided in a tailored
config file. These checks either analyse the agreement with historical reference data or expectations on the scenario data.

# Installation

The package is available via the R package repository of PIK. 

```{r installation}
install.packages("piamValidation", repos = "https://rse.pik-potsdam.de/r/packages")
```

# Usage

wip

If you are unsure where on your computer the output folder is, use ``path.package("piamValidation")`` to find your package.
