---
title: "piamValidation: COMMITED"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  mif: ""
  ref: ""
  cfg: ""
  warning: false
  message: false
  figWidth: 8
---

```{r include=FALSE}
devtools::load_all(".")
library(knitr)
library(ggplot2)
library(ggthemes)
library(plotly)
library(kableExtra)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  #fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)
```

## Import and Prepare Data


```{r}
# Data Preparation
path <- "C:/Users/pascalwe/Data/vs/"
mifs <- c(paste0(path, "REMIND_generic_Bal.mif"),
          paste0(path, "REMIND_generic_NPi.mif"))

hmif <- paste0(path, "historical.mif")

configName <- "validationConfig_COMMITED.csv"

data <- importScenarioData(scenarioPath)

hist <- importReferenceData(referencePath)

cfg <- getConfig(configName) %>%
  cleanConfig()

cfg <- expandPeriods(cfg, data) %>%
  expandVariables(data)

# combine data for each row of the config and bind together
df <- data.frame()
for (i in 1:nrow(cfg)) {
  # TODO: hist should only be needed if category "historical" is in config
  #       validation generally should work without hist data
  df_row <- combineData(data, cfg[i, ], ref_data = hist)
  df <- rbind(df, df_row)
  cat(paste0("Combined row ", i, " of ", nrow(cfg), "\n"))
}

df <- resolveDuplicates(df)

df <- evaluateThresholds(df)

df <- appendTooltips(df)

```

## Heatmaps by Category

### Historic - Relative
Relative deviation to historical reference data. Absolute numbers, so no 
difference is made between being above or below the reference value.
```{r}
c <- "historic"
m <- "relative"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```

### Historic - Difference
Absolute difference to historical reference data
```{r}
c <- "historic"
m <- "difference"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```

### Scenario - Relative
Relative deviation to data point from either:

- period (same scenario/model)
- scenario (same period/model)
- model (same period/scenario - Not supported yet)

```{r}
c <- "scenario"
m <- "relative"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```

### Scenario - Difference
```{r}
c <- "scenario"
m <- "difference"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```

### Scenario - Absolute
```{r}
c <- "scenario"
m <- "absolute"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```


### Scenario - Growthrate
```{r}
c <- "scenario"
m <- "growthrate"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```
