---
title: "piamValidation: REMIND"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  mif: ""
  ref: ""
  cfg: ""
  warning: false
  message: false
  figWidth: 8
---

```{r include=FALSE}
devtools::load_all(".")
library(knitr)
library(ggplot2)
library(ggthemes)
library(plotly)
library(kableExtra)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  #fig.width = params$figWidth,
  message = params$message,
  warning = params$warning
)
```

## Import and Prepare Data


```{r}
# Data Preparation
path <- "C:/Users/pascalwe/Data/vs/"
mifs <- c(paste0(path, "REMIND_generic_Bal.mif"),
          paste0(path, "REMIND_generic_NPi.mif"))

hmif <- paste0(path, "historical.mif")

config <- "validationConfig.csv"

# debugging
# data <- importScenarioData(mifs)
# hist <- importReferenceData(hmif)
# cfg <- getConfig(config) %>% cleanConfig()

df <- validateScenarios(mifs, hmif, config)
df <- appendTooltips(df)
```
## Validation

### Summary
```{r}
# find "critical == yes" data points, that are red/yellow
colors <- c(green  = "#008450",
            yellow = "#EFB700",
            red    = "#B81D13",
            grey   = "#808080")
dt <- dplyr::count(df, critical, check)
kbl(dt) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = F) %>%
  column_spec(2, color = colors[dt$check],
              background = colors[dt$check])
```

## Heatmaps by Category

### Historic - Relative
Relative deviation to historical reference data. Absolute numbers, so no 
difference is made between being above or below the reference value.
```{r}
c <- "historic"
m <- "relative"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```

### Historic - Difference
Absolute difference to historical reference data
```{r}
c <- "historic"
m <- "difference"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```

### Scenario - Relative
Relative deviation to data point from either:

- period (same scenario/model)
- scenario (same period/model)
- model (same period/scenario - Not supported yet)

```{r}
c <- "scenario"
m <- "relative"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```

### Scenario - Difference
```{r}
c <- "scenario"
m <- "difference"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```

### Scenario - Absolute
```{r}
c <- "scenario"
m <- "absolute"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```


### Scenario - Growthrate
```{r}
c <- "scenario"
m <- "growthrate"
d <- df[df$category == c & df$metric == m, ]

if (nrow(d) > 0) {
  vars <- unique(d$variable)
  plot_list <- htmltools::tagList()
    for (i in 1:length(vars)) {
      plot_list[[i]] <- validationHeatmap(d, vars[i], cat = c, met = m)
    }
  plot_list
}
```
