---
title: "REMIND validation v3.5.1"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  mif: ""
  cfg: ""
  extraColors: true
  warning: false
  message: false
  figWidth: 8
---

```{r include=FALSE}
#library(piamValidation)
devtools::load_all()
library(knitr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(plotly)
library(gridExtra)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  message = params$message,
  warning = params$warning
)

config <- "AMT"
```

# Data

Loading data from: `/p/projects/remind/runs/REMIND_2025_06_10/remind`

Using config: `r config` [Link](https://pik-piam.github.io/piamValidation/articles/configs/validationConfig_REMIND.html)

```{r, message = FALSE, warning = FALSE}
# Data Preparation
# r351 <- quitte::as.quitte(
#   c("user/Data/validation_3.5.1/2025-06-10/REMIND_generic_SSP2-NPi2025.mif",
#     "user/Data/validation_3.5.1/2025-06-10/REMIND_generic_SSP2-PkBudg1000.mif",
#     "user/Data/validation_3.5.1/2025-06-10/REMIND_generic_SSP2-PkBudg650.mif",
#     "user/Data/validation_3.5.1/2025-06-10/REMIND_generic_SSP1-PkBudg650.mif"))
# 
# r35 <- quitte::as.quitte(
#   c("user/Data/validation_3.5/v3/REMIND_generic_SSP2-NPi2025.mif",
#     "user/Data/validation_3.5/v3/REMIND_generic_SSP2-PkBudg1000.mif",
#     "user/Data/validation_3.5/v3/REMIND_generic_SSP2-PkBudg650.mif"))
# 
# hist <- quitte::as.quitte("user/Data/validation_3.5.1/historical.mif")
# 
# # average 2020 values over 5 year period to even out Covid shock
# hist_m <- hist %>%
#  filter(period %in% seq(2018, 2022)) %>%
#  magclass::as.magpie(spatial = "region")
# hist_m[, 2020, ] <- magclass::dimSums(hist_m, dim = 2)/5
# hist_smoothed <- quitte::as.quitte(hist_m[, , ]) %>%
#   filter(period == 2020) %>%
#   mutate(model = paste0(model, "_smoothed"))
# 
# hist <- rbind(hist, hist_smoothed)
# 
# # bunkers fix
# #hist$variable <- factor(hist$variable, levels = c(levels(hist$variable), "Emi|CO2|w/o Bunkers|Energy|Demand|Transport"))
# #hist[hist$variable == "Emi|CO2|Energy|Demand|Transport", "variable"] <- "Emi|CO2|w/o Bunkers|Energy|Demand|Transport"
# 
# r351$model <- "REMIND 3.5.1"
# # harmonize scenario names
# # mask <- data.frame(
# #   scenario = c("SSP2-NPi", "SSP2-PkBudg1050", "SSP2-PkBudg650"),
# #   scenario_new = c("SSP2-NPi2025", "SSP2-PkBudg1000","SSP2-PkBudg650")
# # )
# # r351 <- quitte::as.quitte(
# #   quitte::replace_column(r34, mask, scenario = scenario, scenario_new))
# 
# r35$model <- "REMIND 3.5"
# 
# saveRDS(remind2::deletePlus(quitte::as.quitte(rbind(r35, r351, hist))), "data/r351.rds")

data <- readRDS("data/r351.rds")

# for use in lineplots
hist <- data %>% filter(scenario == "historical")

valiData <- validateScenarios(data, config)
valiData <- appendTooltips(valiData)

# reorder regions to World is first
new_order <- unique(intersect(c("World", "GLO",
                                levels(valiData$region)), levels(valiData$region)))
valiData$region <- factor(valiData$region, levels = new_order)
```

# Validation

```{r}
# lineplot wrapper
REMIND_lineplot <- function(var, 
                            reg = "World", 
                            vD = valiData, 
                            df = data, 
                            hist_models = NULL,
                            xlimits = c(2000, 2035)) {

  if (reg == "all") {
    plot_list <- htmltools::tagList()
    for (i in 1:length(unique(valiData$region))) {
  
      reg <- levels(valiData$region)[i]
      line_data <- vD %>%
            filter(variable == var,
                   region == reg)
          
      scen_data <- df %>%
        filter(scenario != "historical",
               variable == var,
               region == reg,
               model %in% c("REMIND 3.5", "REMIND 3.5.1"))
          
      hist_data <- df %>%
        filter(scenario == "historical",
               variable == var,
               region == reg)
    
      if (!is.null(hist_models)) {
        hist_data <- hist_data %>% 
          filter(model %in% hist_models)
      }
      
      plot_list[[i]] <- linePlotThresholds(line_data,
                                           refData = hist_data,
                                           scenData = scen_data,
                                           xlim = xlimits)
    }
    plot_list
  
  } else {
  
    line_data <- vD %>%
      filter(variable == var,
             region == reg)
    
    scen_data <- df %>%
      filter(scenario != "historical",
             variable == var,
             region == reg,
             model %in% c("REMIND 3.5", "REMIND 3.5.1"))
    
    hist_data <- df %>%
      filter(scenario == "historical",
             variable == var,
             region == reg)
    
    if (!is.null(hist_models)) {
      hist_data <- hist_data %>% 
        filter(model %in% hist_models)
    }
    
    linePlotThresholds(line_data,
                       refData = hist_data,
                       scenData = scen_data,
                       xlim = xlim)
    }
  }
```


## Emissions

### Transport {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Emi|CO2|w/o Bunkers|Energy|Demand|Transport"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("CEDS", "EDGARghg", "CEDS_smoothed", "EDGARghg_smoothed"))
```

### Buildings {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Emi|CO2|Energy|Demand|Buildings"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("CEDS", "EDGARghg", "CEDS_smoothed", "EDGARghg_smoothed"))
```

### Energy and IP {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Emi|CO2|Energy and Industrial Processes"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("CEDS", "EDGARghg", "CEDS_smoothed", "EDGARghg_smoothed"))
```

### Energy Supply {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Emi|CO2|Energy|Supply"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("CEDS", "EDGARghg", "CEDS_smoothed", "EDGARghg_smoothed"))
```



## PE

### Coal {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "PE|Coal"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", 
                hist_models = c("IEA-EB-directSum", "BP", "IEA-EB-directSum_smoothed", "BP_smoothed"))
```

### Gas {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "PE|Gas"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", 
                hist_models = c("IEA-EB-directSum", "BP", "IEA-EB-directSum_smoothed", "BP_smoothed"))
```

### Oil {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "PE|Oil"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", 
                hist_models = c("IEA-EB-directSum", "BP", "IEA-EB-directSum_smoothed", "BP_smoothed"))
```


## SE 

### Electricity {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "SE|Electricity"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "BP", "Ember_smoothed", "BP_smoothed"))
```

## FE 

### Total {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "FE"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("IEA-EB-directSum", "IEA-EB-directSum_smoothed"))
```


## Capacities

### Coal {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Coal"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "Ember_smoothed"))
```

### Gas {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Gas"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "Ember_smoothed"))
```

### Solar {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Solar"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "Ember_smoothed"))
```

### Wind {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Wind"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "Ember_smoothed"))
```

### Hydro {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Hydro"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "IRENA", "Ember_smoothed", "IRENA_smoothed"))
```

### Nuclear {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Cap|Electricity|Nuclear"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all", hist_models = c("Ember", "Ember_smoothed"))
```


## Carbon Management

### Storage {.tabset}

#### Heat Map
```{r, fig.width=7, fig.height=6}
var <- "Carbon Management|Storage"
validationHeatmap(valiData[valiData$variable == var, ])
```

#### Line Plots
```{r, fig.width=9, fig.height=5}
REMIND_lineplot(var, reg = "all")
```


# Sustainability Concerns

```{r, fig.width=7, fig.height=6}
var <- "PE|Biomass"
validationHeatmap(valiData[valiData$variable == var, ])
```

## Summary {.tabset}

```{r}
df_summary <- valiData

colors <- c(green     = "#008450",
            yellow    = "#EFB700",
            red       = "#B81D13",
            cyan      = "#66ccee",
            blue      = "#4477aa",
            grey      = "#808080")

# find "critical == yes" data points of each color
summary <- filter(df_summary, region == "World") %>%
  dplyr::count(model, check) %>%
  group_by(model) %>%
  mutate(percent = n / sum(n) * 100)

# change order of colors
summary$check <- factor(
  summary$check, 
  levels = rev(c("green", "yellow", "cyan", "red", "blue", "grey"))
  )

# stacking order and midpoints
summary <- summary %>%
  arrange(model, desc(check)) %>%
  group_by(model) %>%
  mutate(
    cumulative = cumsum(n),
    midpoint = cumulative - (n / 2)
  )


# stacked bar plot in absolute numbers of checks
p1 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(stat = "identity") + # Absolute stacked bars
  scale_fill_manual(values = colors) + # Custom colors
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) +
  geom_text(aes(label = n, y = midpoint), color = "white", size = 3.5) +
  labs(y = "Count", x = "Model") +
  ggtitle("World") + 
  theme(legend.position = "none")

# normalized
p2 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5)) +
  geom_text(aes(label = paste0(round(percent, 1), "%"),
                group = check),
            position = position_fill(vjust = 0.5),
            color = "white",
            size = 3.5) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage", x = "Model") +
  ggtitle("World") + 
  theme(legend.position = "none")


# find "critical == yes" data points of each color
summary <- filter(df_summary, region != "World") %>%
  dplyr::count(model, check) %>%
  group_by(model) %>%
  mutate(percent = n / sum(n) * 100)

# change order of colors
summary$check <- factor(
  summary$check, 
  levels = rev(c("green", "yellow", "cyan", "red", "blue", "grey"))
  )

# stacking order and midpoints
summary <- summary %>%
  arrange(model, desc(check)) %>%
  group_by(model) %>%
  mutate(
    cumulative = cumsum(n),
    midpoint = cumulative - (n / 2)
  )

# stacked bar plot in absolute numbers of checks
p3 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(stat = "identity") + # Absolute stacked bars
  scale_fill_manual(values = colors) + # Custom colors
  theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust = 0.5)) +
  geom_text(aes(label = n, y = midpoint), color = "white", size = 3.5) +
  labs(y = "Count", x = "Model") +
  ggtitle("Regions") + 
  theme(legend.position = "none")

# normalized
p4 <- ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5)) +
  geom_text(aes(label = paste0(round(percent, 1), "%"),
                group = check),
            position = position_fill(vjust = 0.5),
            color = "white",
            size = 3.5) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Percentage", x = "Model") +
  ggtitle("Regions") + 
  theme(legend.position = "none")
```


### Percentage

```{r, fig.width=7, fig.height=4}
grid.arrange(p2, p4, nrow = 1)
```

### Count
```{r, fig.width=7, fig.height=4}
grid.arrange(p1, p3, nrow = 1)
```
