---
title: "piamValidation: Scenario MIP report"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    toc_depth: 4
params:
  mif: ""
  cfg: ""
  extraColors: true
  warning: false
  message: false
  figWidth: 8
---

```{r include=FALSE}
library(piamValidation)
library(knitr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(plotly)
library(madrat)
library(piamInterfaces)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  message = params$message,
  warning = params$warning
)
```

## Import and Prepare Data

Loading data from: 
"scenarios_scenariomip_2024-11-21", "historical_R10_scenarioMIP_2024-11-05.mif"

```{r}
#for (m in params$mif) cat(paste0(normalizePath(m), "\n"))
```

Using config: "scenarioMIP"

```{r}
cat(params$cfg, "\n")
```


```{r, message = FALSE, warning = TRUE}

hist <- quitte::read.snapshot("/historical_R10_scenarioMIP_2024-11-05.mif") # %>%
  #filter(period %in% seq(2018, 2022)) %>%
  #as.magpie(spatial = "region")

# average 2020 values over 5 year period to even out Covid shock
#hist[, 2020, ] <- dimSums(hist, dim = 2)/5
#hist <- quitte::as.quitte(hist[, , ])


#data <- quitte::read.snapshot("/scenarios_scenariomip_2024-11-21.csv")
#saveRDS(data, file = "scenarios_scenariomip_2024-11-21.rds")
data <- readRDS("/scenarios_scenariomip_2024-11-21.rds")

dR10 <- filter(data, region %in% unique(hist$region))

df <- rbind(hist, dR10)

# Data Preparation
df <- validateScenarios(df, "scenarioMIP", extraColors = TRUE)
df <- appendTooltips(df)

# reorder scenario names alphabetically
df$scenario <- factor(df$scenario, levels = sort(levels(df$scenario)))
```

## Validation

### Summary
```{r}
# find "critical == yes" data points of each color
summary <- dplyr::count(df, model, check)

colors <- c(green     = "#008450",
            yellow    = "#EFB700",
            red       = "#B81D13",
            cyan      = "#66ccee",
            blue      = "#4477aa",
            grey      = "#808080")

ggplot(summary, aes(x = model, y= n, fill = check)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = colors) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## Lineplots Historical - Interactive Prototypes

### SSP1

```{r, fig.width=10, fig.height=6}
s <- 1
ssp <- paste0("SSP", s)
valiData <- filter(df, grepl(ssp, scenario),
            metric == "relative", ref_scenario == "historical")
vars <- unique(valiData$variable)

plot_list <- htmltools::tagList()
for (i in 1:length(vars)) {
  var <- as.character(vars[i])
  # this plots only the periods which are being validated 
  # (change in validationConfig)
  p <- linePlotThresholds(valiData, var, reg = "World")
  
  # add historical data
  h <- filter(hist, variable == var, region == "World", 
              model %in% valiData[ ,"ref_model"],
              period %in% seq(2005, 2030)) %>%
    mutate(ref_model = model)
  p <- p +
    geom_point(data = h, aes(x = period, y = value, shape = ref_model),
               size = 1, color = "black")
  
  # add scenario data
  d <- filter(dR10, variable == var, region == "World",
              grepl(ssp, scenario),
              period %in% seq(2005, 2030))
  p <- p +
    geom_line(data = d, aes(x = period, y = value, color = model, linetype = scenario)) +
    ylab(var)
  
  plot_list[[i]] <- plotly::ggplotly(p)
}
plot_list
```


### SSP2

```{r, fig.width=10, fig.height=6}
s <- 2
ssp <- paste0("SSP", s)
valiData <- filter(df, grepl(ssp, scenario),
            metric == "relative", ref_scenario == "historical")
vars <- unique(valiData$variable)

plot_list <- htmltools::tagList()
for (i in 1:length(vars)) {
  var <- as.character(vars[i])
  # this plots only the periods which are being validated 
  # (change in validationConfig)
  p <- linePlotThresholds(valiData, var, reg = "World")
  
  # add historical data
  h <- filter(hist, variable == var, region == "World", 
              model %in% valiData[ ,"ref_model"],
              period %in% seq(2005, 2030)) %>%
    mutate(ref_model = model)
  p <- p +
    geom_point(data = h, aes(x = period, y = value, shape = ref_model),
               size = 1, color = "black")
  
  # add scenario data
  d <- filter(dR10, variable == var, region == "World",
              grepl(ssp, scenario),
              period %in% seq(2005, 2030))
  p <- p +
    geom_line(data = d, aes(x = period, y = value, color = model, linetype = scenario)) +
    ylab(var)
  
  plot_list[[i]] <- plotly::ggplotly(p)
}
plot_list
```

### SSP3

```{r, fig.width=10, fig.height=6}
s <- 3
ssp <- paste0("SSP", s)
valiData <- filter(df, grepl(ssp, scenario),
            metric == "relative", ref_scenario == "historical")
vars <- unique(valiData$variable)

plot_list <- htmltools::tagList()
for (i in 1:length(vars)) {
  var <- as.character(vars[i])
  # this plots only the periods which are being validated 
  # (change in validationConfig)
  p <- linePlotThresholds(valiData, var, reg = "World")
  
  # add historical data
  h <- filter(hist, variable == var, region == "World", 
              model %in% valiData[ ,"ref_model"],
              period %in% seq(2005, 2030)) %>%
    mutate(ref_model = model)
  p <- p +
    geom_point(data = h, aes(x = period, y = value, shape = ref_model),
               size = 1, color = "black")
  
  # add scenario data
  d <- filter(dR10, variable == var, region == "World",
              grepl(ssp, scenario),
              period %in% seq(2005, 2030))
  p <- p +
    geom_line(data = d, aes(x = period, y = value, color = model, linetype = scenario)) +
    ylab(var)
  
  plot_list[[i]] <- plotly::ggplotly(p)
}
plot_list
```

### SSP4

```{r, fig.width=10, fig.height=6}
s <- 4
ssp <- paste0("SSP", s)
valiData <- filter(df, grepl(ssp, scenario),
            metric == "relative", ref_scenario == "historical")
vars <- unique(valiData$variable)

plot_list <- htmltools::tagList()
for (i in 1:length(vars)) {
  var <- as.character(vars[i])
  # this plots only the periods which are being validated 
  # (change in validationConfig)
  p <- linePlotThresholds(valiData, var, reg = "World")
  
  # add historical data
  h <- filter(hist, variable == var, region == "World", 
              model %in% valiData[ ,"ref_model"],
              period %in% seq(2005, 2030)) %>%
    mutate(ref_model = model)
  p <- p +
    geom_point(data = h, aes(x = period, y = value, shape = ref_model),
               size = 1, color = "black")
  
  # add scenario data
  d <- filter(dR10, variable == var, region == "World",
              grepl(ssp, scenario),
              period %in% seq(2005, 2030))
  p <- p +
    geom_line(data = d, aes(x = period, y = value, color = model, linetype = scenario)) +
    ylab(var)
  
  plot_list[[i]] <- plotly::ggplotly(p)
}
plot_list
```


### SSP5

```{r, fig.width=10, fig.height=6}
s <- 5
ssp <- paste0("SSP", s)
valiData <- filter(df, grepl(ssp, scenario),
            metric == "relative", ref_scenario == "historical")
vars <- unique(valiData$variable)

plot_list <- htmltools::tagList()
for (i in 1:length(vars)) {
  var <- as.character(vars[i])
  # this plots only the periods which are being validated 
  # (change in validationConfig)
  p <- linePlotThresholds(valiData, var, reg = "World")
  
  # add historical data
  h <- filter(hist, variable == var, region == "World", 
              model %in% valiData[ ,"ref_model"],
              period %in% seq(2005, 2030)) %>%
    mutate(ref_model = model)
  p <- p +
    geom_point(data = h, aes(x = period, y = value, shape = ref_model),
               size = 1, color = "black")
  
  # add scenario data
  d <- filter(dR10, variable == var, region == "World",
              grepl(ssp, scenario),
              period %in% seq(2005, 2030))
  p <- p +
    geom_line(data = d, aes(x = period, y = value, color = model, linetype = scenario)) +
    ylab(var)
  
  plot_list[[i]] <- plotly::ggplotly(p)
}
plot_list
```


## Historical (non-interactive)

```{r, fig.width=10, fig.height=10}
# all following heat maps should not show historic periods in this document
df <- filter(df, period >= 2020)
```

### All models, all scenarios

Deviation to IEA data (no 2020 smoothing)

```{r, fig.width=10, fig.height=10}
d <- filter(df, metric == "relative", ref_scenario == "historical" | is.na(ref_scenario))
interact <- FALSE

if (interact) {
  # interactive plots can only be printed from outside of loops, use tagList
  # to store html objects in List
  if (nrow(d) > 0) {
    vars <- unique(d$variable)
    plot_list <- htmltools::tagList()
      for (i in 1:length(vars)) {
        plot_list[[i]] <- validationHeatmap(filter(d, variable == vars[i]),
                                            y_plot = "period",
                                            x_plot = "region",
                                            y_facet = "scenario",
                                            x_facet = "model",
                                            interactive = interact) %>% 
          # avoid overlap of title and facet labels (plotly issue)
          layout(title = list(y = .95, xref = "plot"),
                 margin = list(l = 75, t = 150))
      }
      plot_list
  }
} else {
  # non-interactive plots can simply be printed
  vars <- unique(d$variable)
  for (i in 1:length(vars)) {
      p <- validationHeatmap(filter(d, variable == vars[i]),
                              y_plot = "period",
                              x_plot = "region",
                              y_facet = "scenario",
                              x_facet = "model",
                              interactive = interact)
      print(p)
      }
}
```

## Historical (interactive)

### By Model

Deviation to IEA data, averaged from 2018 to 2022 to dampen Covid shock.

```{r, fig.width=10, fig.height=10}
d <- filter(df, metric == "relative", ref_scenario == "historical" | is.na(ref_scenario))
interact <- TRUE

# interactive plots can only be printed from outside of loops, use tagList
# to store html objects in List
if (nrow(d) > 0) {
  models <- unique(d$model)
  plot_list <- htmltools::tagList()
  for (i in 1:length(models)) {
    d_model <- filter(d, model == models[i])
    plot_list[[i]] <- validationHeatmap(d_model, main_dim = "model",
                                        interactive = interact) %>% 
      # avoid overlap of title and facet labels (plotly issue)
      plotly::layout(title = list(y = .95, xref = "plot"),
                     margin = list(l = 0, t = 150, r = 150))
  }
  plot_list
}
```

## Near-Term Trends

### Emissions by Model

Deviation to "Medium" Scenario of the respective SSP in 2030.

```{r, fig.width=10, fig.height=10}
d <- filter(df, metric == "relative", ref_scenario != "historical" | is.na(ref_scenario))
interact <- TRUE

# interactive plots can only be printed from outside of loops, use tagList
# to store html objects in List
if (nrow(d) > 0) {
  models <- unique(d$model)
  plot_list <- htmltools::tagList()
  for (i in 1:length(models)) {
    d_model <- filter(d, model == models[i])
    plot_list[[i]] <- validationHeatmap(d_model, main_dim = "model",
                                        interactive = interact) %>% 
      # avoid overlap of title and facet labels (plotly issue)
      layout(title = list(y = .95, xref = "plot"),
             margin = list(l = 0, t = 150, r = 150))
  }
  plot_list
}
```

### Capacities by Model

Tested against thresholds based on IEA HSMR and IAEA PRIS project pipelines:

https://github.com/pik-piam/mrremind/discussions/540

https://github.com/pik-piam/mrremind/discussions/541

```{r, fig.width=10, fig.height=10}
d <- filter(df, metric == "absolute", grepl("Capacity", variable))
interact <- TRUE

# interactive plots can only be printed from outside of loops, use tagList
# to store html objects in List
if (nrow(d) > 0) {
  models <- unique(d$model)
  plot_list <- htmltools::tagList()
  for (i in 1:length(models)) {
    d_model <- filter(d, model == models[i])
    plot_list[[i]] <- validationHeatmap(d_model, main_dim = "model",
                                        interactive = interact) %>% 
      # avoid overlap of title and facet labels (plotly issue)
      layout(title = list(y = .95, xref = "plot"),
             margin = list(l = 0, t = 150, r = 150))
  }
  plot_list
}
```

### Biomass by Model

```{r, fig.width=10, fig.height=10}
d <- filter(df, metric == "absolute", 
            variable %in% c("Primary Energy|Biomass"))
interact <- TRUE

# interactive plots can only be printed from outside of loops, use tagList
# to store html objects in List
if (nrow(d) > 0) {
  models <- unique(d$model)
  plot_list <- htmltools::tagList()
  for (i in 1:length(models)) {
    d_model <- filter(d, model == models[i])
    plot_list[[i]] <- validationHeatmap(d_model, main_dim = "model",
                                        interactive = interact) %>% 
      # avoid overlap of title and facet labels (plotly issue)
      layout(title = list(y = .95, xref = "plot"),
             margin = list(l = 0, t = 150, r = 150))
  }
  plot_list
}
```

### Carbon Capture by Model

```{r, fig.width=10, fig.height=10}

d <- filter(df, metric == "absolute", 
            variable %in% c("Carbon Capture"))
interact <- TRUE

# interactive plots can only be printed from outside of loops, use tagList
# to store html objects in List
if (nrow(d) > 0) {
  models <- unique(d$model)
  plot_list <- htmltools::tagList()
  for (i in 1:length(models)) {
    d_model <- filter(d, model == models[i])
    plot_list[[i]] <- validationHeatmap(d_model, main_dim = "model",
                                        interactive = interact) %>% 
      # avoid overlap of title and facet labels (plotly issue)
      layout(title = list(y = .95, xref = "plot"),
             margin = list(l = 0, t = 150, r = 150))
  }
  plot_list
}
```
