---
title: "piamValidation: Scenario MIP report"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    toc_depth: 4
params:
  mif: ""
  cfg: ""
  extraColors: true
  warning: false
  message: false
  figWidth: 8
---

```{r, setup, include=FALSE}
library(piamValidation)
library(knitr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(plotly)
library(madrat)
library(piamInterfaces)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  message = params$message,
  warning = params$warning
)

# ensure that the working directory is the project root
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

<!-- ## Import and Prepare Data -->

```{r}
cat(params$cfg, "\n")
```

```{r, message = FALSE, warning = TRUE}
# getwd()

hist <- quitte::read.snapshot("./data/historical_R10_scenarioMIP_2024-11-05.mif")
dR10 <- readRDS("./data/scenarios_scenariomip_2024-11-21_processed.rds") %>%
  filter(region %in% unique(hist$region))
df <- bind_rows(hist, dR10)

files <- c(
  "./data/historical_R10_scenarioMIP_2024-11-05.mif",
  "scenarios_scenariomip_2024-11-21_processed"
)

# Data Preparation
df <- validateScenarios(files, "scenarioMIP", extraColors = TRUE)
df <- appendTooltips(df)

# reorder scenario names alphabetically
df$scenario <- factor(df$scenario, levels = sort(levels(df$scenario)))
```

<!-- ### Naming conventions -->
```{r}
# copy-pasted from REMIND multi-model analysis scripts

plot.emi.scen.letter <- c(
  "High Emissions" = "H",
  "Medium Emissions" = "M",
  "Medium-Low Emissions" = "ML",
  "Low Emissions" = "L",
  "Low Overshoot" = "LOS",
  "Very Low Emissions" = "VL"
)

plot.emi.scen.colors <- c(
  "High Emissions" = "darkred",
  "Medium Emissions" = "#FF6F6F",
  "Medium-Low Emissions" = "#B282AE",
  "Low Emissions" = "#6495ED",
  "Low Overshoot" = "#FF8C00",
  "Very Low Emissions" = "#6B8E23"
)

plot.model.colors <- c(
  # scenario models
  "AIM" = "#4D5CAD",
  "COFFEE" = "#69BA7F",
  "GCAM" = "#759EA8",
  "IMAGE" = "#868367",
  "MESSAGEix-GLOBIOM" = "#892F71",
  "REMIND-MAgPIE" = "#facb1e",
  "WITCH" = "#fb6a4a"
)

plot.model.letters <- c(
  "AIM" = "A",
  "COFFEE" = "C",
  "MESSAGEix-GLOBIOM" = "M",
  "GCAM" = "G",
  "REMIND-MAgPIE" = "R",
  "IMAGE" = "I",
  "WITCH" = "W"
)

plot.SSP.linetype <- c(
    "SSP1" = "dashed",
    "SSP2" = "solid",
    "SSP3" = "dotted",
    "SSP4" = "longdash",
    "SSP5" = "dotdash"
  )

```


## Validation

### Data sources

#### 2020 vetting

Deviation to IEA data.
For 2020, the tolerance range includes both the 2020 value and the 2018-2022 average.

#### Near-term emissions

Deviation to "Medium" Scenario of the respective SSP in 2030.


#### Near-term capacities

Tested against thresholds based on IEA HSMR and IAEA PRIS project pipelines:

<https://github.com/pik-piam/mrremind/discussions/540>

<https://github.com/pik-piam/mrremind/discussions/541>

### Summary

```{r}
# find "critical == yes" data points of each color
summary <- dplyr::count(df, model, check) %>%
  mutate(model = substr(model, 1, 5))
  #revalue.levels(model = plot.model.letters)

ggplot(summary, aes(x = model, y = n, fill = check)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = validationHeatmap.colors) +
  theme(axis.text.x = element_text(angle = 0, vjust = 0, hjust=0.5))
```

<!-- ## Historical vetting by model -->
```{r, fig.width=10, fig.height=6}

interactiveLinplots_model <- function(modelArg = "AIM") {
  valiData <- df %>%
    #revalue.levels(emi.scen = plot.emi.scen.letter) %>%
    filter(model == modelArg,
           metric == "relative",
           ref_scenario == "historical")
  
  vars <- unique(valiData$variable)
  
  local_hist <- hist %>%
      filter(variable %in% vars,
             region == "World",
             model %in% valiData$ref_model,
             period %in% seq(2005, 2030)) %>%
      mutate(ref_model = model)
  
  local_scen <- dR10 %>%
      #revalue.levels(emi.scen = plot.emi.scen.letter) %>%
      filter(variable %in% vars,
             region == "World",
             model == modelArg,
             period %in% seq(2005, 2030)) %>%
      mutate(scenario = paste0(" ", ssp, gsub("Default|Variant","",def.var)))
  
  plot_list <- htmltools::tagList()
  for (i in 1:length(vars)) {
    var <- as.character(vars[i])
    
    # this plots only the periods which are being validated 
    # (change in validationConfig)
    p <- linePlotThresholds(valiData, var, reg = "World", refData = T)
    
    # add historical data
    h <- local_hist %>% filter(variable == var)
    p <- p +
      geom_point(data = h, aes(x = period, y = value), #, shape = ref_model),
                 size = 1, color = "black")
    
    # add scenario data
    d <- local_scen %>% filter(variable == var)
    p <- p +
      geom_line(data = d, aes(x = period, y = value, color = emi.scen, linetype = scenario, group = ssp)) +
      scale_color_manual(values = plot.emi.scen.colors)
      #labs(color = "Scenario") +
      #scale_linetype_manual(values = plot.SSP.linetype) +

    plot_list[[i]] <- plotly::ggplotly(p)
  }
  return(plot_list)
}

```
<!-- ### AIM -->
<!-- ```{r, fig.width=10, fig.height=6} -->
<!-- interactiveLinplots_model("REMIND-MAgPIE") -->
<!-- ``` -->


## Historical vetting by scenario narrative
```{r, fig.width=10, fig.height=6}

interactiveLinplots_emiScen <- function(emiScen = "VL") {
  valiData <- df %>%
    revalue.levels(emi.scen = plot.emi.scen.letter) %>%
    filter(emi.scen == emiScen,
           metric == "relative",
           ref_scenario == "historical")
  
  vars <- unique(valiData$variable)
  
  local_hist <- hist %>%
      filter(variable %in% vars,
             region == "World",
             model %in% valiData$ref_model,
             period %in% seq(2005, 2030)) %>%
      mutate(ref_model = model)
  
  local_scen <- dR10 %>%
      revalue.levels(emi.scen = plot.emi.scen.letter) %>%
      filter(variable %in% vars,
             region == "World",
             emi.scen == emiScen,
             period %in% seq(2005, 2030)) %>%
      mutate(scenario = paste0(" ", ssp, gsub("Default|Variant","",def.var)))
  
  plot_list <- htmltools::tagList()
  for (i in 1:length(vars)) {
    var <- as.character(vars[i])
    
    # this plots only the periods which are being validated 
    # (change in validationConfig)
    p <- linePlotThresholds(valiData, var, reg = "World", refData = T)
    
    # add historical data
    h <- local_hist %>% filter(variable == var)
    p <- p +
      geom_point(data = h, aes(x = period, y = value), #, shape = ref_model),
                 size = 1, color = "black")
    
    # add scenario data
    d <- local_scen %>% filter(variable == var)
    p <- p +
      geom_line(data = d, aes(x = period, y = value, color = model, linetype = scenario, group = ssp)) +
      #labs(color = "Scenario") +
      #scale_linetype_manual(values = plot.SSP.linetype) +
      scale_color_manual(values = plot.model.colors)
    
    plot_list[[i]] <- plotly::ggplotly(p)
  }
  return(plot_list)
}

```

### VL
```{r, fig.width=10, fig.height=6}
interactiveLinplots_emiScen("VL")
```

### LOS
```{r, fig.width=10, fig.height=6}
interactiveLinplots_emiScen("LOS")
```

### L
```{r, fig.width=10, fig.height=6}
interactiveLinplots_emiScen("L")
```

### ML
```{r, fig.width=10, fig.height=6}
interactiveLinplots_emiScen("ML")
```

### M
```{r, fig.width=10, fig.height=6}
interactiveLinplots_emiScen("M")
```

### H
```{r, fig.width=10, fig.height=6}
interactiveLinplots_emiScen("H")
```


<!-- ## Historical vetting by SSP -->

<!-- ```{r, fig.width=10, fig.height=6} -->

<!-- interactiveLinplots <- function(sspArg = "SSP1") { -->
<!--   valiData <- df %>% -->
<!--     filter(ssp == sspArg, -->
<!--            metric == "relative", -->
<!--            ref_scenario == "historical") -->

<!--   vars <- unique(valiData$variable) -->

<!--   local_hist <- hist %>% -->
<!--       filter(variable %in% vars, -->
<!--              region == "World", -->
<!--              model %in% valiData$ref_model, -->
<!--              period %in% seq(2005, 2030)) %>% -->
<!--       mutate(ref_model = model) -->

<!--   local_scen <- dR10 %>% -->
<!--       filter(variable %in% vars, -->
<!--              region == "World", -->
<!--              ssp == sspArg, -->
<!--              period %in% seq(2005, 2030)) %>% -->
<!--       revalue.levels(emi.scen = plot.emi.scen.letter) %>% -->
<!--       mutate(scenario = paste0(" ", emi.scen, gsub("Default|Variant","",def.var))) -->


<!--   plot_list <- htmltools::tagList() -->
<!--   for (i in 1:length(vars)) { -->
<!--     var <- as.character(vars[i]) -->

<!--     # this plots only the periods which are being validated  -->
<!--     # (change in validationConfig) -->
<!--     p <- linePlotThresholds(valiData, var, reg = "World", refData = T) -->

<!--     # add historical data -->
<!--     h <- local_hist %>% filter(variable == var) -->
<!--     p <- p + -->
<!--       geom_point(data = h, aes(x = period, y = value), #, shape = ref_model), -->
<!--                  size = 1, color = "black") -->

<!--     # add scenario data -->
<!--     d <- local_scen %>% filter(variable == var) -->
<!--     p <- p + -->
<!--       geom_line(data = d, aes(x = period, y = value, color = model, linetype = scenario)) -->

<!--     plot_list[[i]] <- plotly::ggplotly(p) -->
<!--   } -->
<!--   return(plot_list) -->
<!-- } -->

<!-- ``` -->

<!-- ### SSP1 -->
<!-- ```{r, fig.width=10, fig.height=6} -->
<!-- interactiveLinplots("SSP1") -->
<!-- ``` -->


<!-- ### SSP2 -->
<!-- ```{r, fig.width=10, fig.height=6} -->
<!-- interactiveLinplots("SSP2") -->
<!-- ``` -->

<!-- ### SSP3 -->
<!-- ```{r, fig.width=10, fig.height=6} -->
<!-- interactiveLinplots("SSP3") -->
<!-- ``` -->

<!-- ### SSP4 -->
<!-- ```{r, fig.width=10, fig.height=6} -->
<!-- interactiveLinplots("SSP4") -->
<!-- ``` -->

<!-- ### SSP5 -->
<!-- ```{r, fig.width=10, fig.height=6} -->
<!-- interactiveLinplots("SSP5") -->
<!-- ``` -->

<!-- ## 2020 overview (non-interactive) -->

```{r, fig.width=10, fig.height=10}
# all following heat maps should not show historic periods in this document
df_heatmap <- df %>%
  filter(period >= 2020)

# when there are several checks for the same variable, display only the more extreme color
# minor issue: if only the max is the violated, hoverbox won't show the min
# issue: if a strict rule is overwritten by a lax rule, we may get unwanted color warnings
df_heatmap <- df_heatmap %>%
  mutate(checkRank = check) %>%
  revalue.levels(checkRank = validationHeatmap.colorRanks) %>%
  group_by(region,variable,unit,period, model, scenario, value, ssp,emi.scen,def.var) %>%
  filter(checkRank == max(checkRank)) %>%
  select(-checkRank)

# testing example:
# ddd <- df_heatmap %>%
#   filter(model == "IMAGE", region == "India+ (R10)", variable == "Final Energy", emi.scen == "Very Low Emissions", ssp == "SSP1", period == 2020) %>%
#   mutate(checkRank = check) %>%
#   revalue.levels(checkRank = validationHeatmap.colorRanks) %>%
#   group_by(region,variable,unit,period, model, scenario, value, ssp,emi.scen,def.var) %>%
#   filter(checkRank == max(checkRank))

```

<!-- Deviation to IEA data (with 2020 smoothing) -->

<!-- ```{r, fig.width=10, fig.height=10} -->
<!-- d <- df_heatmap %>% -->
<!--   filter(metric == "relative", ref_scenario == "historical" | is.na(ref_scenario)) -->
<!-- interact <- F -->

<!-- if (interact) { -->
<!--   # interactive plots can only be printed from outside of loops, use tagList -->
<!--   # to store html objects in List -->
<!--   if (nrow(d) > 0) { -->
<!--     vars <- unique(d$variable) -->
<!--     plot_list <- htmltools::tagList() -->
<!--       for (i in 1:length(vars)) { -->
<!--         plot_list[[i]] <- validationHeatmap(filter(d, variable == vars[i]), -->
<!--                                             y_plot = "period", -->
<!--                                             x_plot = "region", -->
<!--                                             y_facet = "scenario", -->
<!--                                             x_facet = "model", -->
<!--                                             interactive = interact) %>%  -->
<!--           # avoid overlap of title and facet labels (plotly issue) -->
<!--           layout(title = list(y = .95, xref = "plot"), -->
<!--                  margin = list(l = 75, t = 150)) -->
<!--       } -->
<!--       plot_list -->
<!--   } -->
<!-- } else { -->
<!--   # non-interactive plots can simply be printed -->
<!--   vars <- unique(d$variable) -->
<!--   for (i in 1:length(vars)) { -->
<!--       p <- validationHeatmap(filter(d, variable == vars[i]), -->
<!--                               y_plot = "period", -->
<!--                               x_plot = "region", -->
<!--                               y_facet = "scenario", -->
<!--                               x_facet = "model", -->
<!--                               interactive = interact) -->
<!--       print(p) -->
<!--       } -->
<!-- } -->
<!-- ``` -->

<!-- ## 2020 by model (interactive) -->

<!-- Deviation to IEA data. -->
<!-- For 2020, the tolerance range includes both the 2020 value and the 2018-2022 average. -->

```{r, modelHeatmap, fig.width=10, fig.height=10}
modelHeatmap <- function(d) {
  d <- d %>%
    filter(metric == "relative", ref_scenario == "historical" | is.na(ref_scenario))
  interact <- TRUE
  
  # interactive plots can only be printed from outside of loops, use tagList
  # to store html objects in List
  if (nrow(d) > 0) {
    models <- unique(d$model)
    plot_list <- htmltools::tagList()
    for (i in 1:length(models)) {
      plot_list[[i]] <- d %>%
        filter(model == models[i]) %>%
        validationHeatmap(main_dim = "model",
                          interactive = interact,
                          titleSuffix = " 2020") %>% 
        # avoid overlap of title and facet labels (plotly issue)
        plotly::layout(title = list(y = .95, xref = "plot"),
                       margin = list(l = 0, t = 150, r = 150))
    }
    return(plot_list)
  }
}
```

```{r}
#modelHeatmap(df_heatmap)
```


<!-- ## Definitions -->

<!-- ### 2020 vetting -->

<!-- Deviation to IEA data. -->
<!-- For 2020, the tolerance range includes both the 2020 value and the 2018-2022 average. -->


<!-- ### Near-term emissions -->

<!-- Deviation to "Medium" Scenario of the respective SSP in 2030. -->

```{r, modelEmissions, fig.width=10, fig.height=10}
modelEmissions <- function(d) {
  d <- d %>%
    filter(metric == "relative", ref_scenario != "historical" | is.na(ref_scenario))
  interact <- TRUE
  
  # interactive plots can only be printed from outside of loops, use tagList
  # to store html objects in List
  if (nrow(d) > 0) {
    models <- unique(d$model)
    plot_list <- htmltools::tagList()
    for (i in 1:length(models)) {
      plot_list[[i]] <- d %>%
        filter(model == models[i]) %>%
        validationHeatmap(main_dim = "model",
                          interactive = interact,
                          titleSuffix = "near-term emissions") %>% 
        # avoid overlap of title and facet labels (plotly issue)
        layout(title = list(y = .95, xref = "plot"),
               margin = list(l = 0, t = 150, r = 150))
    }
    return(plot_list)
  }
}
```

```{r}
#modelEmissions(df_heatmap)
```


<!-- ### Near-term capacities -->

<!-- Tested against thresholds based on IEA HSMR and IAEA PRIS project pipelines: -->

<!-- <https://github.com/pik-piam/mrremind/discussions/540> -->

<!-- <https://github.com/pik-piam/mrremind/discussions/541> -->

```{r, modelCapacities, fig.width=10, fig.height=10}
modelCapacities <- function(d) {
  d <- d %>%
    filter(metric == "absolute", grepl("Capacity", variable))
  interact <- TRUE
  
  # interactive plots can only be printed from outside of loops, use tagList
  # to store html objects in List
  if (nrow(d) > 0) {
    models <- unique(d$model)
    plot_list <- htmltools::tagList()
    for (i in 1:length(models)) {
      plot_list[[i]] <- d %>%
        filter(model == models[i]) %>%
        validationHeatmap(main_dim = "model",
                          interactive = interact,
                          titleSuffix = "near-term capacities") %>% 
        # avoid overlap of title and facet labels (plotly issue)
        layout(title = list(y = .95, xref = "plot"),
               margin = list(l = 0, t = 150, r = 150))
    }
    return(plot_list)
  }
}
```

```{r}
#modelCapacities(df_heatmap)
```


<!-- ### Near-term biomass -->

```{r, modelBiomass, fig.width=10, fig.height=10}
modelBiomass <- function(d) {
  d <- d %>%
    filter(metric == "absolute",
           variable %in% c("Primary Energy|Biomass"))
  interact <- TRUE
  
  # interactive plots can only be printed from outside of loops, use tagList
  # to store html objects in List
  if (nrow(d) > 0) {
    models <- unique(d$model)
    plot_list <- htmltools::tagList()
    for (i in 1:length(models)) {
      plot_list[[i]] <- d %>%
        filter(model == models[i]) %>%
        validationHeatmap(main_dim = "model",
                          interactive = interact,
                          titleSuffix = "near-term biomass") %>% 
        # avoid overlap of title and facet labels (plotly issue)
        layout(title = list(y = .95, xref = "plot"),
               margin = list(l = 0, t = 150, r = 150))
    }
    return(plot_list)
  }
}
```

```{r}
#modelBiomass(df_heatmap)
```


<!-- ### Near-term carbon capture -->

```{r, modelCC, fig.width=10, fig.height=10}

modelCC <- function(d, title = TRUE) {
  d <- d %>%
    filter(metric == "absolute", 
           variable %in% c("Carbon Capture"))
  interact <- TRUE
  
  # interactive plots can only be printed from outside of loops, use tagList
  # to store html objects in List
  if (nrow(d) > 0) {
    models <- unique(d$model)
    plot_list <- htmltools::tagList()
    for (i in 1:length(models)) {
      plot_list[[i]] <- d %>%
        filter(model == models[i]) %>%
        validationHeatmap(main_dim = "model",
                          interactive = interact,
                          titleSuffix = "near-term carbon capture") %>% 
        # avoid overlap of title and facet labels (plotly issue)
        layout(title = list(y = .95, xref = "plot"),
               margin = list(l = 0, t = 150, r = 150))
    }
    return(plot_list)
  }
}
```

```{r}
#modelCC(df_heatmap)
```

<!-- ## Results by model -->

```{r}
# this chunk prints Rmd lines that we copy paste below for the HTML rendering
messageChunk <- function(...) {
  message("```{r}")
  message(...)
  message("```")
}
for (modelLoop in unique(df$model) %>% sort) {
  #cat("  \n## Model ", modelLoop)
  message("## Model ", modelLoop)
  messageChunk("modelLoop='", modelLoop, "'
d <- df_heatmap %>% filter(model==modelLoop)
interactiveLinplots_model(modelLoop)")
  messageChunk("modelHeatmap(d)")
  messageChunk("modelCapacities(d)")
  messageChunk("modelEmissions(d)")
  messageChunk("modelBiomass(d)")
  messageChunk("modelCC(d)")
}
```


## Model AIM
```{r}
modelLoop='AIM'
d <- df_heatmap %>% filter(model==modelLoop)
interactiveLinplots_model(modelLoop)
```
```{r}
modelHeatmap(d)
```
```{r}
modelCapacities(d)
```
```{r}
modelEmissions(d)
```
```{r}
modelBiomass(d)
```
```{r}
modelCC(d)
```
## Model COFFEE
```{r}
modelLoop='COFFEE'
d <- df_heatmap %>% filter(model==modelLoop)
interactiveLinplots_model(modelLoop)
```
```{r}
modelHeatmap(d)
```
```{r}
modelCapacities(d)
```
```{r}
modelEmissions(d)
```
```{r}
modelBiomass(d)
```
```{r}
modelCC(d)
```
## Model GCAM
```{r}
modelLoop='GCAM'
d <- df_heatmap %>% filter(model==modelLoop)
interactiveLinplots_model(modelLoop)
```
```{r}
modelHeatmap(d)
```
```{r}
modelCapacities(d)
```
```{r}
modelEmissions(d)
```
```{r}
modelBiomass(d)
```
```{r}
modelCC(d)
```
## Model IMAGE
```{r}
modelLoop='IMAGE'
d <- df_heatmap %>% filter(model==modelLoop)
interactiveLinplots_model(modelLoop)
```
```{r}
modelHeatmap(d)
```
```{r}
modelCapacities(d)
```
```{r}
modelEmissions(d)
```
```{r}
modelBiomass(d)
```
```{r}
modelCC(d)
```
## Model MESSAGEix-GLOBIOM
```{r}
modelLoop='MESSAGEix-GLOBIOM'
d <- df_heatmap %>% filter(model==modelLoop)
interactiveLinplots_model(modelLoop)
```
```{r}
modelHeatmap(d)
```
```{r}
modelCapacities(d)
```
```{r}
modelEmissions(d)
```
```{r}
modelBiomass(d)
```
```{r}
modelCC(d)
```
## Model REMIND-MAgPIE
```{r}
modelLoop='REMIND-MAgPIE'
d <- df_heatmap %>% filter(model==modelLoop)
interactiveLinplots_model(modelLoop)
```
```{r}
modelHeatmap(d)
```
```{r}
modelCapacities(d)
```
```{r}
modelEmissions(d)
```
```{r}
modelBiomass(d)
```
```{r}
modelCC(d)
```
## Model WITCH
```{r}
modelLoop='WITCH'
d <- df_heatmap %>% filter(model==modelLoop)
interactiveLinplots_model(modelLoop)
```
```{r}
modelHeatmap(d)
```
```{r}
modelCapacities(d)
```
```{r}
modelEmissions(d)
```
```{r}
modelBiomass(d)
```
```{r}
modelCC(d)
```
