---
title: "piamValidation: default report"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  mif: ""
  cfg: ""
  extraColors: true
  warning: false
  message: false
  figWidth: 8
---

```{r include=FALSE}
library(piamValidation)
library(knitr)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(plotly)

knitr::opts_chunk$set(
  echo = FALSE,
  error = TRUE,
  message = params$message,
  warning = params$warning
)
```

## Import and Prepare Data

Loading data from:

```{r}
for (m in params$mif) cat(paste0(normalizePath(m), "\n"))
```

Using config:

```{r}
cat(params$cfg, "\n")
```


```{r, message = FALSE, warning = TRUE}
# Data Preparation
df <- validateScenarios(params$mif, params$cfg, extraColors = params$extraColors)
df <- appendTooltips(df)
```

## Validation

### Summary
```{r}
# find "critical == yes" data points of each color
dplyr::count(df, critical, check)
```

## Heatmaps by Category

### Historic - Relative
Relative deviation to historical reference data.
```{r}
interact = TRUE

# plot one heat map for each variable in this category
d <- filter(df, metric == "relative", ref_scenario == "historical")
vars <- unique(d$variable)

if (nrow(d) > 0) {
  if (interact) {  # interactive plots need to be collected in a tagList
    plot_list <- htmltools::tagList()
      for (i in 1:length(vars)) {
        plot_data <- filter(d, variable == vars[i])
        plot_list[[i]] <- validationHeatmap(plot_data)
      }
    plot_list
  } else {  # non-interactive plots can simply be printed in a loop
    for (i in 1:length(vars)) {
        p <- validationHeatmap(d, vars[i], interactive = interact)
        print(p)
    }
  }
}
```

### Historic - Difference
Absolute difference to historical reference data
```{r}
interact = TRUE

# plot one heat map for each variable in this category
d <- filter(df, metric == "difference", ref_scenario == "historical")
vars <- unique(d$variable)

if (nrow(d) > 0) {
  if (interact) {  # interactive plots need to be collected in a tagList
    plot_list <- htmltools::tagList()
      for (i in 1:length(vars)) {
        plot_data <- filter(d, variable == vars[i])
        plot_list[[i]] <- validationHeatmap(plot_data)
      }
    plot_list
  } else {  # non-interactive plots can simply be printed in a loop
    for (i in 1:length(vars)) {
        p <- validationHeatmap(d, vars[i], interactive = interact)
        print(p)
    }
  }
}
```

### Scenario - Relative
Relative deviation to data point from either:

- period (same scenario/model)
- scenario (same period/model)
- model (same period/scenario)

```{r}
interact = TRUE

# plot one heat map for each variable in this category
d <- filter(df, metric == "relative", ref_scenario != "historical" | is.na(ref_scenario))
vars <- unique(d$variable)

if (nrow(d) > 0) {
  if (interact) {  # interactive plots need to be collected in a tagList
    plot_list <- htmltools::tagList()
      for (i in 1:length(vars)) {
        plot_data <- filter(d, variable == vars[i])
        plot_list[[i]] <- validationHeatmap(plot_data)
      }
    plot_list
  } else {  # non-interactive plots can simply be printed in a loop
    for (i in 1:length(vars)) {
        p <- validationHeatmap(d, vars[i], interactive = interact)
        print(p)
    }
  }
}
```

### Scenario - Difference
```{r}
interact = TRUE

# plot one heat map for each variable in this category
d <- filter(df, metric == "difference", ref_scenario != "historical" | is.na(ref_scenario))
vars <- unique(d$variable)

if (nrow(d) > 0) {
  if (interact) {  # interactive plots need to be collected in a tagList
    plot_list <- htmltools::tagList()
      for (i in 1:length(vars)) {
        plot_data <- filter(d, variable == vars[i])
        plot_list[[i]] <- validationHeatmap(plot_data)
      }
    plot_list
  } else {  # non-interactive plots can simply be printed in a loop
    for (i in 1:length(vars)) {
        p <- validationHeatmap(d, vars[i], interactive = interact)
        print(p)
    }
  }
}
```

### Scenario - Absolute
```{r}
interact = TRUE

# plot one heat map for each variable in this category
d <- filter(df, metric == "absolute")
vars <- unique(d$variable)

if (nrow(d) > 0) {
  if (interact) {  # interactive plots need to be collected in a tagList
    plot_list <- htmltools::tagList()
      for (i in 1:length(vars)) {
        plot_data <- filter(d, variable == vars[i])
        plot_list[[i]] <- validationHeatmap(plot_data)
      }
    plot_list
  } else {  # non-interactive plots can simply be printed in a loop
    for (i in 1:length(vars)) {
        p <- validationHeatmap(d, vars[i], interactive = interact)
        print(p)
    }
  }
}
```


### Scenario - Growthrate
```{r}
interact = TRUE

# plot one heat map for each variable in this category
d <- filter(df, metric == "growthrate")
vars <- unique(d$variable)

if (nrow(d) > 0) {
  if (interact) {  # interactive plots need to be collected in a tagList
    plot_list <- htmltools::tagList()
      for (i in 1:length(vars)) {
        plot_data <- filter(d, variable == vars[i])
        plot_list[[i]] <- validationHeatmap(plot_data)
      }
    plot_list
  } else {  # non-interactive plots can simply be printed in a loop
    for (i in 1:length(vars)) {
        p <- validationHeatmap(d, vars[i], interactive = interact)
        print(p)
    }
  }
}
```
